---
title: "Advantage"
author: "Orion Bowers"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  html_document: default
---

```{r setup}
set.seed(101)
num_trials <- 10000000

roll_die <- function(n, sides) {
  if (sides == 20) {
    floor(runif(n, 1, sides+1))
  } else if (sides==10) {
    floor(runif(n, 0, sides))
  } else if (sides==100) {
    floor(runif(n, 0, sides))
  }
}
  
sim_rolls <- function(n, sides, bonus=0, penalty=0, push=FALSE) {
    if (!(sides == 20 || sides == 10 || sides == 100)) {
        return("Error: Please use a d20, d10 or d100")
    }
    if (bonus < 0 || bonus > 2) {
      return("Error: Please restrict bonus dice from 0 to 2")
    }
    if (penalty < 0 || penalty > 2) {
      return("Error: Please restrict penalty dice from 0 to 2")
    }
    if (bonus != 0 && penalty != 0) {
      return("Error: Choose either bonus or penalty dice (subtract one from the other if needed)")
    }
  
    if (sides == 20) {
      # d10
      result <- roll_die(n, sides)
      if (bonus > 0) {
        rolls <- replicate(bonus, roll_die(n, sides), simplify = FALSE) # Bonus dice rolled
        result <- do.call(pmax, c(list(result), rolls)) # Calls pmax on result and bonus rolls
      } else if (penalty > 0) {
        rolls <- replicate(penalty, roll_die(n, sides), simplify = FALSE) # Penalty dice rolled
        result <- do.call(pmin, c(list(result), rolls)) # Calls pmin on result and penalty rolls
      }
      
    } else if (sides == 10) {
      # d10*10 + 1d10 (only reroll the tens die with bonus/penalty)
      tens <- roll_die(n, sides)*10
      ones <- roll_die(n, sides)

      if (bonus > 0) {
        rolls <- replicate(bonus, roll_die(n, sides)*10, simplify = FALSE) # Bonus dice rolled
        tens <- do.call(pmin, c(list(tens), rolls)) # Calls pmax on result and bonus rolls
      } else if (penalty > 0) {
        rolls <- replicate(penalty, roll_die(n, sides)*10, simplify = FALSE) # Penalty dice rolled
        tens <- do.call(pmax, c(list(tens), rolls)) # Calls pmin on result and penalty rolls
      }
      
      result <- tens+ones
    } else {
      # 1d100
      result <- roll_die(n, sides)
      if (bonus > 0) {
        rolls <- replicate(bonus, roll_die(n, sides), simplify = FALSE) # Bonus dice rolled
        result <- do.call(pmin, c(list(result), rolls)) # Calls pmax on result and bonus rolls
      } else if (penalty > 0) {
        rolls <- replicate(penalty, roll_die(n, sides), simplify = FALSE) # Penalty dice rolled
        result <- do.call(pmax, c(list(result), rolls)) # Calls pmin on result and penalty rolls
      }
    }
  
  return(result)
    
}
```



## Introduction

Differences between systems.

Advantage/Disadvantage.

Pushing/Rerolling

Cutoffs on 1/4, 1/2, 3/4 of a typical character (8-16 and 1% - 99%)

## D&D

```{r}
standard <- sim_rolls(num_trials, 20)
advantage <- sim_rolls(num_trials, 20, bonus = 1)
advantage2 <- sim_rolls(num_trials, 20, bonus = 2)
disadvantage <- sim_rolls(num_trials, 20, penalty = 1)
disadvantage2 <- sim_rolls(num_trials, 20, penalty = 2)

plot(table(standard), main='Standard Roll')

plot(table(advantage), main= 'Advantage')
plot(table(advantage2), main='Double Advantage')

plot(table(disadvantage), main= 'Disadvantage')
plot(table(disadvantage2), main='Double Disadvantage')

```




## Call of Cthulhu

### Side Note

1d10*10 + 1d10 isn't quite the same as 1d100, but it's very similar in practice and is probably not worth the hassle of using a full d100 while playing. The difference is most evident when using bonus and penalty dice because the probabilities for each number don't follow a smooth curve when using 2d10s.

```{r}
standard <- sim_rolls(num_trials, 10)
bonus <- sim_rolls(num_trials, 10, bonus = 1)
bonus2 <- sim_rolls(num_trials, 10, bonus = 2)
penalty <- sim_rolls(num_trials, 10, penalty = 1)
penalty2 <- sim_rolls(num_trials, 10, penalty = 2)

plot(table(standard), main = "Standard Roll 1d10*10+1d10")

plot(table(bonus), main = "1 Bonus Die")
plot(table(bonus2), main = "2 Bonus Dice")

plot(table(penalty), main = "1 Penalty Die")
plot(table(penalty2), main = "2 Penalty Dice")
```


```{r}
standard <- sim_rolls(num_trials, 100)
bonus <- sim_rolls(num_trials, 100, bonus = 1)
bonus2 <- sim_rolls(num_trials, 100, bonus = 2)
penalty <- sim_rolls(num_trials, 100, penalty = 1)
penalty2 <- sim_rolls(num_trials, 100, penalty = 2)

plot(table(standard), main = "Standard Roll 1d100")

plot(table(bonus), main = "1 Bonus Die")
plot(table(bonus2), main = "2 Bonus Dice")

plot(table(penalty), main = "1 Penalty Die")
plot(table(penalty2), main = "2 Penalty Dice")
```

```{r}
low_boundary <- 25
mid_boundary <- 50
up_boundary <- 75
```